---
title: "Claudia RMarkdown Presentation"
format: beamer
output:
  ioslides_presentation: default
  beamer_presentation:
    theme: sky
    slide_level: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
## Goal

- Development of a Non-Lethal Genetic Technique for Determining Age of Pacific Halibut

NOTE: Aging is typically conducted via the extraction of a fishes otolith. Which ultimately ends up contributing largely towards species mortality rates -- since you have to bonk (x_x) the fish

## Initial Methods

- Fin clips are collected through fishery independent set-line surveys (FISS). Each sample has been double aged already by the International Pacific Halibut Commission (IPHC)

- DNA gets extracted and pooled into libraries containing only 48 samples. Done in batches to achieve 20x coverage with each pool

- Code gets written for trimming, alignments, methylation extraction, and more for initial clean-up and quality checks 

- Once this is complete I will start the modeling  process and identification of any CpG sites  

## Slide with Baseline R Code

```{r, engine='bash', eval=FALSE, echo = TRUE}

Using project sample code as an example. 
This is an example of what I type into the server for TrimGalore

trim_galore --output_dir trimgalore-out --fastqc_args \
"--outdir fastqc-out" \
  --basename ${SAMPLE2} --non_directional --rrbs --paired \
  --2colour=20 --keep \
  umi-add/${SAMPLE2}_R1.fq.gz umi-add/${SAMPLE2}_R2.fq.gz
```

## Initial Output

```{r, engine='bash', eval=FALSE, echo = TRUE}
## add code with document output
plot()
```

